---
title: "Tutorial Section 2 Exercise 2.3 & 2.4"
author: "Josiah"
date: "Sunday, May 03, 2015"
output: html_document
---

### Synopisis
This is my attempt at doing the exercises 2.3 & 2.4 for section 2 of [An Introduction To Data Cleaning With R](http://cran.r-project.org/doc/contrib/de_Jonge+van_der_Loo-Introduction_to_data_cleaning_with_R.pdf) article by **Edwin de Jonge** and **Mark van der Loo**. This is by no means the correct answers as they were not peer reviewed or presented to the authors of the document.

### Some functions that are later used
This is my function to get max length of elements in a list
```{r}
#Maximum length of a list
maxlength <- function(x){
  if(!is.list(x) | is.null(x)){
    return("Either not a list or a null element")
  }
  rslt = 0;
  for(i in 1:length(x)){
    if(length(x[[i]]) > rslt){
      rslt <- length(x[[i]])
    }
  }
  return(rslt)
}
```
This is my assignfields function inspired by the one provided by the article
```{r}
assignfields <- function(x, n = 3){
  if(!is.vector(x) | is.null(x) | n <= 0){
    return("Either not a vector, a null element or your
           pivot index is not at least 1")
  }
  out <- character(n)
  # get names
  i <- grepl("[[:alpha:]]",x)
  out[1] <- x[i]
  # get age
  # Since weight is decimal, Hoping that none is written as
  # integer or 12.0 for example, the following might work
  i <- which(as.integer(x) == x)
  out[2] <- ifelse(length(i) > 0, x[i], NA)
  # get weight
  if(!any(grepl(",", x))){
    i <- which(as.integer(x) != x)
  } else {
    i <- which(grepl(",", x))
  }
  out[3] <- ifelse(length(i) >0, x[i], NA)
  out
}
```

## Exercise 2.3 : Using readLines
1 Getting the data
```{r}
fileurl <- "http://raw.githubusercontent.com/edwindj/datacleaning/master/Rnw/syllabus/files/example.txt"
download.file(fileurl, destfile = "example.txt", method = "curl")
```
2 Reading data into object
```{r}
(Example <- readLines("example.txt"))
```
3 Using grepl to separate comments & data
```{r}
(com_vec <- Example[grepl("^//", Example)])
(dat_vec <- Example[!grepl("^//", Example)])
```
4 Extract the date from the first comment line
```{r}
(date_created <- gsub("^ ", "", strsplit(com_vec[1],":")[[1]][2]))
```
5 Read data into matrix

* Split data by ';'
```{r}
(my_list <- strsplit(dat_vec, ";"))
```
* Find Maximum length of elements of list and append shorter elements with NA

The maximum number of fields returned by split is `r maxlength(my_list)`
```{r, warning=FALSE}
(dat_fields <- lapply(my_list, assignfields))
```
* Unlisting and turning into a matrix
```{r}
(my_mat <- matrix(
    unlist(dat_fields)
   , nrow=length(dat_fields)
   , byrow=TRUE))
```
6 Extract names of the fields and set them as column names
```{r}
# Notice that each comment line can be split on the : mark. Once you split it on that mark, you need to pick the second element of each vector in the list. A clever way to do that is unlisting and targeting the even indexes
Names <- strsplit(com_vec[2:4], ":")
# The rep(...) targets the even indexes: 2,4,6
(Names <- gsub("^ ", "", unlist(Names)[rep(c(FALSE,TRUE),2)]))
colnames(my_mat) <- Names
my_mat
```

## Exercise 2.4 : Coercing mat to data.frame
1 Coerce to data frame : avoid factors
```{r}
(my_dat <- data.frame(my_mat, stringsAsFactors = F))
sapply(my_dat, class)
```
2 Use a string distance technique to create factor variable
```{r}
codes <- c("male", "female")
dist_df <- adist(my_dat$Gender, codes)
colnames(dist_df) <- codes
rownames(dist_df) <- my_dat$Gender
dist_df
ind <- apply(dist_df, 1, which.min)
(coded_dat <- data.frame(gender = my_dat$Gender, coded = codes[ind]))
tmp_vec <- plyr::revalue(coded_dat$coded, c(male = "man", female = "woman"))
(my_dat$Gender <- tmp_vec)
my_dat
sapply(my_dat, class)
```
3 Coerce Age into Integers
```{r}
(my_dat$Age..in.years. <- as.integer(my_dat$Age))
sapply(my_dat, class)
```
4 Coerce Weight into numeric
```{r}
my_dat$Weight..in.kg. <- gsub(",", ".", my_dat$Weight)
my_dat$Weight..in.kg. <- as.numeric(my_dat$Weight..in.kg.)
my_dat
sapply(my_dat, class)
```
